---
title: "Combining GPS & Demo Features"
author: "JJC, updated by Claire Punturieri"
date: "`r lubridate::today()`"
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

## Notes
- remove aud id?


## Set up

```{r}
#| include: false

library(tidyverse)
options(conflicts.policy = "depends.ok")

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
source("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("~/github/proj_risk/shared/fun_rename.R")
path_shared <- format_path("risk/data_processed/shared")
path_gps <- format_path("risk/data_processed/gps")
```

Read in feature files
```{r}
gps <- read_csv(here::here(path_shared, "features_gps_day_24h.csv"), 
                 show_col_types = FALSE) 
demos <- read_csv(here::here(path_shared, "features_demographics.csv"), 
                  show_col_types = FALSE)
weather <- read_csv(here::here(path_shared, "features_weather.csv"), 
                 show_col_types = FALSE)
strat <- read_csv(here::here(path_shared, "strat_lh.csv"), 
                    show_col_types = FALSE) 
```

Check rows
```{r}
nrow(gps)
nrow(demos)
nrow(weather)
nrow(strat)
```

Verify expected # of subids

- GPS sample  for GPS, weather
- Full sample for demos
```{r}
gps |> distinct(subid) |> nrow() 
demos |> distinct(subid) |> nrow()
weather |> distinct(subid) |> nrow() 
strat |> distinct(subid) |> nrow() 
```

Remove duplicated rows because of Daylight savings
Only needed for 24 hour rolls
```{r}
gps <- gps |> distinct(subid, dttm_label, .keep_all = TRUE)
```

Clean feature names with John's function.
```{r}
gps <- gps |> 
  rename_gps_features_stage1() |>
  reorder_feature_names(type = "gps")
```


Temporary cleaning of feature names, maybe integrate with John's code or just change how features are generated.

> add in to jjc rename function

```{r}
# change weather names
weather <- weather |>
  rename_with(function(x) {

    # remove suffixes
    x <- sub("(\\.snow|\\.pcpn)$", "", x)

    # main transformation
    x <- gsub(
      "^p(\\d+)\\.l0\\.(r|d)([a-z]+)_([a-z0-9\\.]+)$",
      "gps_\\4.\\3\\2.p\\1",
      x
    )

    # convert r/d to raw/diff
    x <- sub("(\\.[a-z]+)r(\\.p\\d+)$", "\\1raw\\2", x)
    x <- sub("(\\.[a-z]+)d(\\.p\\d+)$", "\\1diff\\2", x)

    # special cases
    x <- sub("^p(\\d+)\\.l0\\.rratecount\\.([a-z0-9_]+)$", 
             "gps_\\2.ratecountraw.p\\1", x)
    
    x <- sub("^p(\\d+)\\.l0\\.dratecount\\.([a-z0-9_]+)$", 
             "gps_\\2.ratecountdiff.p\\1", x)

    x
  })

```

Append

```{r}
gps <- gps |>
  rename_with(~ case_when(
    str_remove(.x, "^gps_") %in% c("subid", "dttm_label", "lapse") ~ .x,
    str_starts(.x, "gps_") ~ .x,
    str_starts(.x, "act_") ~ str_replace(.x, "^act_", "gps_"),
    TRUE ~ str_c("gps_", .x)
  ))
```


Look at features for each signal

```{r}
names(gps)
names(demos)
names(weather)
names(strat)
```


Join all files using left joins starting with GPS for that sample
Also join in strat lh
```{r}
all <- left_join(gps, demos, by = c("subid"))
nrow(all)

all <- left_join(all, strat, by = c("subid"))
nrow(all)

all <- left_join(all, weather, by = c("subid", "dttm_label", "lapse"))
nrow(all)
```

Check outcome.  Looks good!
```{r}
all |> tab(lapse) 
```

Filtering decisions.
```{r}
all <- all |> 
  select(-contains("dif")) |> 
  select(-contains("p6"), -contains("p12"), -contains("p48"), -contains("p72")) |> 
  select(-contains("other")) |> 
  select(-contains("_cm_"), -contains("transit")) |> 
  select(-contains("himed"), -contains("lowno"))
```

Write out file
```{r}
all |>
  select(subid, dttm_label, lapse,
         starts_with("id_"), everything()) |>
  relocate(strat, .after = last_col()) |>
  relocate(starts_with("demo_"), .after = lapse) |>
  write_csv(here::here(path_gps, 
                            str_c("features_combined.csv")))
```
