---
title: "Make combined features for GPS study"
author: "Claire Punturieri"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
params:
  version: 1
---

# Housekeeping

## Code Status

Complete as of 6/2025.

## Notes

Add in circadian movement data once file is complete.

# Set up

## Set up environment

```{r}
#| message: false
#| warning: false

options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
library(tidyverse)
```

## Paths

```{r}
path_shared <- format_path("risk/data_processed/shared")
path_gps <- format_path("risk/data_processed/gps")
```

## Packages and source

```{r}
#| message: false
#| warning: false
#| echo: false

# for data wrangling
library(tidyverse)
library(lubridate)

# helpful lab functions
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")

# adjust plot visuals
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max = Inf)
```

## Load in data

Load in GPS features.
```{r}
gps <- read_csv(here::here(path_shared, "features_gps_day_1h.csv"), 
                 show_col_types = FALSE)
```

Load in weather features.
```{r}
weather <- read_csv(here::here(path_shared, "features_weather.csv"), 
                 show_col_types = FALSE)
```

Load in circadian movement features.
```{r}
# placeholder
```

Load in stratification data, merge relevant column.
```{r}
strat <- read_csv(here::here(path_shared, "strat_lh.csv"), 
                 show_col_types = FALSE)
```

## Join data

```{r}
features <- gps |> 
  left_join(weather, by = c("subid", "dttm_label", "lapse")) |> 
  left_join(strat, by = "subid")
```

## EDA

Verify percentage of lapses looks reasonable.
```{r}
features |> 
  janitor::tabyl(lapse)
```

Check names of all features.
```{r}
features |> colnames()
```

Overview across all categories.
```{r}
features |>  
  skimr::skim_without_charts() |> 
  select(skim_variable, n_missing, complete_rate, numeric.p0, numeric.p50, 
         numeric.p100) |>
  filter(!skim_variable == "subid" & !skim_variable == "label_num") |> 
  arrange(desc(numeric.p100), skim_variable) 
```

## Save out
```{r}
features |>
  write_csv(here::here(path_gps, str_c("features_combined.csv")))
```


