---
title: "Baseline Testing"
author: "Claire Punturieri"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
params:
  window: "day"
  roll_dur: 24 # 1 24
  sample: "gps" # ema messages
---

## Notes

This script is designed to test out the calculation of baseline ratesum. The problems that have been identified are: 1) raw and difference scores are highly correlated; and 2) difference scores appear to not go below 0 (but theoretically they SHOULD be bounded between -1 and 1).

## To do

- Early on study, late on study
- Common places (like home) related to uncommon places (like medium risk locations)

## Set Up

### Set up environment basics

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

options(conflicts.policy = "depends.ok")
```

### Paths

```{r}
source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

name_job <- str_c("features_", params$sample, "_", params$window, "_", params$roll_dur, "h")

path_chtc <- format_path(str_c("risk/chtc/shared/features_gps/", name_job, "/output"))
path_shared <- format_path("risk/data_processed/shared")
features_file <- str_c("batch_features.csv")
```

### Additional packages and source

```{r}
#| message: false
#| warning: false
#| echo: false

# helpful lab functions
source("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")

# adjust plot visuals
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max = Inf)
```

### Load in features, demo issues

Import features file.
```{r}
features <- read_csv(here::here(path_chtc, features_file), 
                 show_col_types = FALSE) |>
  filter(subid == 10)
```

Demo high correlations.
```{r}
cor(features$p24.l0.rratesum_duration.type.home, features$p24.l0.dratesum_duration.type.home, use = "pairwise.complete.obs")
```

Demo bounding of difference features.
```{r}
features |>  
  skimr::skim_without_charts() |> 
  select(skim_variable, numeric.p0, numeric.p50, 
         numeric.p100) |>
  filter(!skim_variable == "subid" & !skim_variable == "label_num") |> 
  filter(grepl("dratesum", skim_variable))
```

## GPS Ratesum Function Step-Through

Load in fun_features.R file here.


Load in needed files for testing. This set-up parallels what is done in the mak_features_gps.R script.
```{r}
dist_max <- 0.031 # tolerance threshold for how far someone can be for know location (~=50m)
window <- "day"  #window for calculating labels
roll_dur <- 24 # 1
sample <- "gps"

data <- #read_csv("gps.csv.xz", show_col_types = FALSE) |>
  # for testing:
  read_csv(file.choose(), show_col_types = FALSE) |>
  # variable conversions
  mutate(time = with_tz(time, tz = "America/Chicago"),
         dist = dist / 1609.344,
         duration = duration / 60,
         speed = dist / duration,
         dist_context = dist_context / 1609.344) |>
  rename(dttm_obs = time) |>
  mutate(duration = if_else(dist > 0.01 & duration == 0, NA_real_, duration),
         duration = if_else(speed > 100, NA_real_, duration),
         duration = if_else(duration > .5 & dist > 0.31, NA_real_, duration),
         known_loc = if_else(dist_context <= dist_max & speed <= 4, "yes", "no"),
         known_loc = if_else(is.na(known_loc), "no", known_loc),
         transit = if_else(speed <= 4, "no", "yes"),
         home = if_else(type == "home", "yes", "no"),
         evening_out = if_else((as.numeric(hour(dttm_obs)) >= 19 | as.numeric(hour(dttm_obs)) < 4)
                               & home == "no",
                               "yes", "no"),
         pass_adi_cats = case_when(pass_adi_state_decile %in% c(1,2) ~ "1",
                              pass_adi_state_decile %in% c(3,4) ~ "2",
                              pass_adi_state_decile %in% c(5,6) ~ "3",
                              pass_adi_state_decile %in% c(7,8) ~ "4",
                              pass_adi_state_decile %in% c(9,10) ~ "5",
                              TRUE ~ NA_character_)
         ) |>
  filter(subid == 10)

labels <- read_csv(file.choose(),
                   show_col_types = FALSE) |>
  mutate(dttm_label = with_tz(day_start, tz = "America/Chicago"),
         day_end = with_tz(day_end, tz = "America/Chicago")) |> 
  filter(subid == 10)

dates <- data |> 
  arrange(subid, dttm_obs) |> 
  group_by(subid) |> slice(1) |> 
  rename(data_start = dttm_obs) |> 
  select(subid, data_start)

# Initialize period durations and lead hours ------------------
period_durations <- c(6, 12, 24, 48, 72, 168)
lead <- 0
```

Set testing constants.
```{r}
x_all <- data
data_start <- dates
col_name <- "duration"
subid <- 10
the_subid <- 10
period_duration <- 168
```

### Testing common location (home) early on study

Set basics for testing.
```{r}
label <- labels |> slice(4)
dttm_label <-  label$dttm_label
the_dttm_label <- dttm_label
context_col_name <- "type"
context_value <- "home"
```

Begin stepping through function.

Print the following as you go through for checking.
```{r}
true_period_duration / 24 # number of days in window, contingent on amount of data that exists, will be ~4 here

x_pd |> view() # will be four days of data, as this is early on study

x |> select(known_loc) |> view() # x filters down to known locations, check this is true

base_duration / 24 # for days on study -- will be same as true_period_duration if early on study

baseline # RATE of time spent at home since on study

raw # RATE of time spent at home in past week
```

Raw and baseline will be the SAME in this instance because it is early on study. Therefore, the difference will be 0.

### Testing common location (home) late on study

```{r}
label <- labels |> slice(85)
dttm_label <-  label$dttm_label
the_dttm_label <- dttm_label
context_col_name <- "type"
context_value <- "home"
```

Print the following as you go through for checking.
```{r}
true_period_duration / 24 # correct pd if time on study is not longer than window, should be 7 here

x_pd |> view() # should be one week, because now we're later on study

x |> select(known_loc) |> view() # x filters down to known locations, check this is true

base_duration / 24 # for days on study

baseline # RATE of time spent at home since on study

raw # RATE of time spent at home in past week
```

The difference score here would still be positive.

Let's look at how baseline is calculated. To calculate our ratesum we will be dividing
```{r}
baseline <- get_x_period(the_subid, the_dttm_label, x_all = (x_all |> filter(known_loc == "yes") |> filter(type == "home")), lead, period_duration = Inf) |>
          correct_last_duration(the_dttm_label) |>
          summarise("base" := ratesum(.data[[col_name]], base_duration)) |>
          pull(base) |> 
          clip_ratesum()

baseline2 <- get_x_period(the_subid, the_dttm_label, x_all = x, lead, period_duration = Inf) |>
          correct_last_duration(the_dttm_label) |>
          summarise("base" := ratesum(.data[[col_name]], base_duration)) |>
          pull(base) |> 
          clip_ratesum()
```

### Testing uncommon location (medium risk) early on study
```{r}
label <- labels |> slice(4)
dttm_label <-  label$dttm_label
the_dttm_label <- dttm_label
context_col_name <- "risk"
context_value <- "medium"
```


### Testing uncommon location (medium risk) late on study
```{r}
label <- labels |> slice(85)
dttm_label <-  label$dttm_label
the_dttm_label <- dttm_label
context_col_name <- "risk"
context_value <- "medium"
```
