---
title: "Make Study Dates for GPS"
author: "John Curtin for EMA, updated for GPS by Claire Punturieri"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

# Housekeeping

## Code Status

Complete

This script creates a GPS study-level dates file that includes the study start date, which is also called data_start (see below).  It also includes study end and ema_end for use in mak_labels.  

This script is essentially the same as Kendra's meta script and the EMA study-level script because the GPS study currently includes the same sample as the meta study (further exclusion of subjects based on completeness of GPS data is done elsewhere). 


## Conclusions   

- John and Kendra have decided to drop subid 104's data for the following reasons:   
  - They had lapses every day on study except one day.    
  - Only had 75 surveys where a lapse was not reported.   
  - Viewing their lapse times it appears they were drinking constantly (morning and 
  night).   
  - They consistently report being uncertain that their goal is to be abstinent 
  (uncertain on 125 of 137 lapses. They also report they are uncertain in this goal 
  at followup 1 and 2.    
  - They are ultimately discontinued since they were struggling to gain sobriety.   
  - Unfortunately this drops 109 valid lapses.    


- John and Kendra have decided to drop subid 269's data for the following reasons:       
  - They completed 10-15 surveys on many days on study (avg number of surveys per 
  day is 6.76).  
  - Their responses indicate careless responding - they were filling 2-4 surveys out 
  within an hour of each other and the responses to the questions were completely different.     
  - They have questionable no lapse labels - they reported no lapses while on study but 
  according to notes left two messages for study staff where they admitted to drinking 
  over the weekend.   
  

- John and Kendra have decided to drop subid 204's data for the following reasons:    
  - Subid 204 had extremely poor compliance. 33 out of 89 study days had an EMA completed. 
  They only did a total of 5 surveys between followup 2 and 3.    
  - We don't trust their lapse labels - They report several lapses during their interviews 
  but times appear questionable (same time every night). They only report 1 lapse with EMA.
  - From notes - "Participant did not do many surveys during their second month of participation. 
  At their Follow-up 2 visit they reported several lapses that were not documented in their 
  EMAs - estimated lapse days/times in subid's raw data log."  
  - JC note: "There are issues with 204. They are missing lapses reported by interview. But they  
  also stopped doing any ema by 5/17 even though their study end date was 6/13. Probably need to 
  drop them for lapse analyses for anything after 5/17.  Probably also need to add in their 
  reported lapses at follow-up 2. OR we drop them at the end of follow-up 1 or wherever their 
  ema gets sketchy"    


- John and Kendra have decided to decided to retain 128's data even though they have over 100 lapses for 
the following reasons:   
  - Compliance is good (averaged over 3 surveys per day, no days without an EMA).       
  - completed the study for the full 90 days.    
  - appeared to still want abstinence as they reported they were uncertain to ema_1_5 
  on only 3 surveys. They reported they were uncertain that their goal was to remain 
  abstinent at followup 1 and confirmed their goal was to remain abstinent at followup 2.    
  - Has more non-lapse surveys than lapse surveys.   
  

- All final timezones are in America/Chicago timezone.    


## Set up

## Environment
```{r}
#| message: false
#| warning: false

options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true", 
                      sha1 = "a58e57da996d1b70bb9a5b58241325d6fd78890f")
```

### Paths
```{r}
path_processed <- format_path("studydata/risk/data_processed/shared")
path_ema <- format_path("studydata/risk/data_processed/ema")
path_gps <- format_path("studydata/risk/data_processed/gps")
```

## Packages and plot settings
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)

# adjust plot visuals
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max = Inf)
```

## Cross-reference study visit dates with EMA

### Visit/EMA Dates

Read in visit dates file for all subjects, but filter to only include start date, follow-up 1 date, and end date.
```{r}
dates <- read_csv(here::here(path_processed, "visit_dates.csv"), show_col_types = FALSE) |> 
  select(subid, study_start = start_study, study_end = end_study, followup_1) |>  
  mutate(study_start = force_tz(study_start, tzone = "America/Chicago"),
         study_end = force_tz(study_end, tzone = "America/Chicago")) |>  
  glimpse()
```

Filter out subids who did not make it to follow-up 1.
```{r}
dates <- dates |>  
  filter(!is.na(followup_1)) |>  
  select(-followup_1) |>  
  glimpse()
```

Filter out excluded subids for lapse reporting issues.
```{r}
dates <- dates |>  
  filter(!subid %in% c(104, 269, 204)) |> 
  glimpse()
```

Filter out excluded subids for problems with GPS.
```{r}
dates <- dates |>  
  filter(!subid %in% c(2, 21, 51, 56, 190)) |>  
  glimpse()
```

Adjust start start and end dates for missing GPS.  Notes in mak_gps_enriched.qmd
```{r}
dates$study_start[dates$subid == 19] = as_datetime("2017-11-26", tz = "America/Chicago")
dates$study_end[dates$subid == 19] = as_datetime("2018-01-17", tz = "America/Chicago")

dates$study_start[dates$subid == 28] = as_datetime("2017-12-18", tz = "America/Chicago")

dates$study_end[dates$subid == 30] = as_datetime("2017-12-24", tz = "America/Chicago")

dates$study_end[dates$subid == 34] = as_datetime("2018-03-01", tz = "America/Chicago")

#dates$study_end[dates$subid == 39] = as_datetime("2018-05-11", tz = "America/Chicago")
dates$study_start[dates$subid == 39] = as_datetime("2018-03-02", tz = "America/Chicago")

#dates$study_end[dates$subid == 40] = as_datetime("2018-04-28", tz = "America/Chicago")
dates$study_end[dates$subid == 40] = as_datetime("2018-04-29", tz = "America/Chicago")

dates$study_start[dates$subid == 42] = as_datetime("2018-04-11", tz = "America/Chicago")

dates$study_end[dates$subid == 47] = as_datetime("2018-04-16", tz = "America/Chicago")

dates$study_end[dates$subid == 63] = as_datetime("2018-06-07", tz = "America/Chicago")

dates$study_start[dates$subid == 65] = as_datetime("2018-07-18", tz = "America/Chicago")

## 66 has a period of 8 missing days - to discuss
# added by cp
dates$study_start[dates$subid == 66] = as_datetime("2018-07-03", tz = "America/Chicago")

dates$study_start[dates$subid == 74] = as_datetime("2018-07-19", tz = "America/Chicago")

#dates$study_end[dates$subid == 76] = as_datetime("2018-08-28", tz = "America/Chicago")
dates$study_end[dates$subid == 76] = as_datetime("2018-08-20", tz = "America/Chicago")

dates$study_end[dates$subid == 78] = as_datetime("2018-08-28", tz = "America/Chicago")

dates$study_end[dates$subid == 80] = as_datetime("2018-07-31", tz = "America/Chicago")

dates$study_end[dates$subid == 84] = as_datetime("2018-08-06", tz = "America/Chicago")

dates$study_start[dates$subid == 117] = as_datetime("2018-10-14", tz = "America/Chicago")

# added by cp
dates$study_start[dates$subid == 134] = as_datetime("2018-09-28", tz = "America/Chicago")

# added by cp
dates$study_end[dates$subid == 137] = as_datetime("2018-12-13", tz = "America/Chicago")

# 139 has period of 11 missing days - to discuss
# added by cp
dates$study_start[dates$subid == 139] = as_datetime("2018-11-19", tz = "America/Chicago")


# removed by cp - not sure if we should be removing any of this person's data
#dates$study_start[dates$subid == 161] = as_datetime("2018-12-20", tz = "America/Chicago")

dates$study_start[dates$subid == 167] = as_datetime("2019-02-06", tz = "America/Chicago")

# removed by cp - not sure if we should be removing any of this person's data
#dates$study_start[dates$subid == 169] = as_datetime("2019-02-19", tz = "America/Chicago")

# 171 has 1 observation per day for last ~month of study - to discuss

dates$study_start[dates$subid == 178] = as_datetime("2019-02-25", tz = "America/Chicago")

# removed by cp - not sure if we should be removing any of this person's data
#dates$study_start[dates$subid == 180] = as_datetime("2019-03-02", tz = "America/Chicago")

# 185 has many 1 observation days - to discuss

dates$study_end[dates$subid == 203] = as_datetime("2019-05-19", tz = "America/Chicago")

dates$study_end[dates$subid == 207] = as_datetime("2019-05-19", tz = "America/Chicago")

dates$study_start[dates$subid == 214] = as_datetime("2019-04-17", tz = "America/Chicago")

dates$study_end[dates$subid == 218] = as_datetime("2019-06-04", tz = "America/Chicago")

# 241 missing 1 week of data
dates$study_start[dates$subid == 241] = as_datetime("2019-09-12", tz = "America/Chicago")

# cp note: even within this salvaged gap there is a 6 day period of missing data - to discuss
dates$study_end[dates$subid == 265] = as_datetime("2019-10-07", tz = "America/Chicago")

# added by cp
dates$study_start[dates$subid == 270] = as_datetime("2019-09-19", tz = "America/Chicago")
```


### Data Start Date

Add column for data_start - will be = study start for EMA b/c everyone started EMA
with study start. Keeping this new column because we use data_start for other studies
and some study generic code may expect it.
```{r}
dates <- dates |>  
  mutate(data_start = study_start) |> 
  glimpse()
```


### EMA end date

Get last EMA data for EMA end.
```{r}
ema <- read_csv(here::here(path_processed, "ema_morning.csv"), show_col_types = FALSE) |>  
    select(subid, start_date)  |>   
  bind_rows(read_csv(here::here(path_processed, "ema_later.csv"), show_col_types = FALSE) |>  
    select(subid, start_date)) |> 
  mutate(subid = as.numeric(subid),
         start_date = with_tz(start_date, tzone = "America/Chicago")) |>  
  rename(ema_end = start_date) |> 
  arrange(subid, desc(ema_end)) |> 
  group_by(subid) |>  
  slice(1) |>  
  glimpse()
```


Add `ema_end` date.
```{r}
dates <- dates |>  
  left_join(ema, by = "subid") |>  
  relocate(subid, study_start, data_start) |>  
  glimpse()
```

### GPS start and end dates

Get first and last GPS date for GPS start and end.
```{r}
gps <- read_csv(here::here(path_processed, "gps.csv"), show_col_types = FALSE) |>  
  group_by(subid) |> 
  arrange(time) |> 
  slice(1, n()) |>
  mutate(start_gps = first(time),
         end_gps = last(time)) |> 
  select(subid, start_gps, end_gps) |> 
  distinct()
```

Add `gps_start` and `gps_end` date.
```{r}
dates <- dates |>  
  left_join(gps, by = "subid") |>  
  relocate(subid, study_start, data_start) |>  
  glimpse()
```

## Save

Save dates.
```{r}
dates  |>  
  write_csv(here::here(path_gps, "study_dates.csv")) |>  
  glimpse()
```

