---
title: "Make Study Dates for GPS"
author: "John Curtin for EMA, updated for GPS by Claire Punturieri"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

# Housekeeping

## Code Status

Complete

This script creates a GPS study-level dates file that includes the study start date, which is also called data_start (see below).  It also includes study end and ema_end for use in mak_labels.  

This script is essentially the same as Kendra's meta script and the EMA study-level script because the GPS study currently includes the same sample as the meta study (further exclusion of subjects based on completeness of GPS data is done elsewhere). 


## Conclusions

### EMA

- John and Kendra have decided to drop subid 104's data for the following reasons:   
  - They had lapses every day on study except one day.    
  - Only had 75 surveys where a lapse was not reported.   
  - Viewing their lapse times it appears they were drinking constantly (morning and 
  night).   
  - They consistently report being uncertain that their goal is to be abstinent 
  (uncertain on 125 of 137 lapses. They also report they are uncertain in this goal 
  at followup 1 and 2.    
  - They are ultimately discontinued since they were struggling to gain sobriety.   
  - Unfortunately this drops 109 valid lapses.    


- John and Kendra have decided to drop subid 269's data for the following reasons:       
  - They completed 10-15 surveys on many days on study (avg number of surveys per 
  day is 6.76).  
  - Their responses indicate careless responding - they were filling 2-4 surveys out 
  within an hour of each other and the responses to the questions were completely different.     
  - They have questionable no lapse labels - they reported no lapses while on study but 
  according to notes left two messages for study staff where they admitted to drinking 
  over the weekend.   
  

- John and Kendra have decided to drop subid 204's data for the following reasons:    
  - Subid 204 had extremely poor compliance. 33 out of 89 study days had an EMA completed. 
  They only did a total of 5 surveys between followup 2 and 3.    
  - We don't trust their lapse labels - They report several lapses during their interviews 
  but times appear questionable (same time every night). They only report 1 lapse with EMA.
  - From notes - "Participant did not do many surveys during their second month of participation. 
  At their Follow-up 2 visit they reported several lapses that were not documented in their 
  EMAs - estimated lapse days/times in subid's raw data log."  
  - JC note: "There are issues with 204. They are missing lapses reported by interview. But they  
  also stopped doing any ema by 5/17 even though their study end date was 6/13. Probably need to 
  drop them for lapse analyses for anything after 5/17.  Probably also need to add in their 
  reported lapses at follow-up 2. OR we drop them at the end of follow-up 1 or wherever their 
  ema gets sketchy"    


- John and Kendra have decided to decided to retain 128's data even though they have over 100 lapses for 
the following reasons:   
  - Compliance is good (averaged over 3 surveys per day, no days without an EMA).       
  - completed the study for the full 90 days.    
  - appeared to still want abstinence as they reported they were uncertain to ema_1_5 
  on only 3 surveys. They reported they were uncertain that their goal was to remain 
  abstinent at followup 1 and confirmed their goal was to remain abstinent at followup 2.    
  - Has more non-lapse surveys than lapse surveys.   

### GPS

*Removal*

- John and Claire have decided to drop 2, 21, 51, 56, 190 due to insufficient GPS data.
  - 2 had technical issues (likely related to Samsung phone) and is missing 94% of data [5 lapses]
  - 21 had technical issues (likely related to Samsung phone) and is missing 56% of data with large gaps [no lapses]
  - 51 had technical issues (likely related to Samsung phone) is missing 74% of data [14 lapses]
  - 56 missing 92% of data [26 lapses]
  - 190 had technical issues throughout study; despite troubleshooting, only a few days of data were
  able to be collected [no lapses]

*GPS problems that require adjustment of study dates*

- 19 had tech issues (samsung) with many gaps in gps.  Mostly continuous block between
11/26 - 1/17.  Adjust study start to 11/26 and study end to 1/17 [ lost no lapses and retained 2 lapses]
- 28 missing data for most of the first month.   Data are present afterward.  Adjust study start to 12/18
- 30 has good data through 12/24.   set study end to 12/24
- 34 is missing data right near end.  Set study end to 3/1.  [Lost one valid lapse but retained MANY]
- 39 has gap at beginning, set study start to 3/2.
- 40 has big gap start  on 4/30.  Set study end to 4/29
- 42 had many big gaps throughout their data. Mostly continuous block started at 4/11 to end of study
- 47 has almost no data after 4/16.  Set study end to 4/16 [lost two lapses but have many more]
- 65 had tech (samsung) issues with many gaps in gps.  Has completed data starting at 7/18.
Set study start to 7/18 [lost two lapses]
- 74 had problems with Moves.  Gaps in data using Moves.  Switched to FollowMee on 7/19, good
data quality. Set study start to 7/19
- 76 has last date of GPS on 8/20.  Set study end to 8/20
- 78 has last day of GPS on 8/28.  Set study end to 8/28
- 80 has good data for about first month (through 7/31).  No data after that b/c app not on phone.  
set study end to 7/31 (lost 6 valid lapses but retains 1 lapse)
to resolve issue.   Set study start to 7/19
- 84 had many phone issues (lost phone, missing data, out of country with GPS off).  But had good data through 8/6.
Set study end to 8/6 [lost 1 lapse and retained none].
- 117 turned off location services repeatedly b/c of data plan issues. Many gaps. There is a somewhat gap free period started at 10/14 - end of study.  Set study start to 10/14 [loses 3 but retains 7 lapses]
Set study start to 4/11.  [lost 4 but retained 14 lapses]
- 134 missing most data first ~week on study, set start to 9/28.
- 167 had gps problems in first month.  Resolved at FU1.  Recode study start to 2/6 (data of FU1 when GPS starts working)
- 169 has five day gap at study start.  Set study start to 2/19 [lost one lapse but retained MANY more]
- 171 has one observation per day starting on 3/21, set study end to 3/20.
- 178 has one big gap in middle of study 2/11 - 2/25.  Data is continuous from 2/25 through 
end of study.  Set study start to 2/25.  [lost 5 but retained 2 lapses]
- 180 traveled out of country for most of first month.  Set start date to 3/2 
[lost many invalid and a couple of valid lapses. Many remain]
- 207 had issues with broken phone, missing days, and out of country travel. Despite this, data were ok (some gaps but not really big ones) through 5/19.   
End study on 5/19 (lost two lapses)
- 214 missing big gap (4/5 - 4/16) near start of study because of tech issue.   
Set study start to 4/17 [lost 4 lapses but has many more]
- 218 has last day GPS on 6/4.  End study on that day [lost 3 lapses]
- 265 has OK data from start through 10/7.  But there is one 6 day gap. Set study end to 10/7
- 270 has inconsistent data for first ~10 days, set study start to 9/19


*Data OK despite missing; No adjustments*

- 3 has missing days but these appear to result from not leaving home for long periods do to a fractured leg.  Retain all data
- 26 has some gaps.  One due to being housebound and others but data seem sufficient 
once gaps are handled.
- 27 has some missing days but OK
- 32 has some gaps but OK
- 53 has some gaps but data appear usable as is
- 63 has some big gaps Also stays home a lot. Could consider setting study end to 6/7,
but would mean losing 1 lapse and retaining 0 lapses
- 66 has a few gaps but OK
- 79 has some missing days but OK
- 137 missing one five day gap but OK
- 139 has gaps but rarely left home.  Likely OK
- 161 has big gap from 12/10 - 12/19. Could consider setting study start to 12/20
- 185 OK.  Missing days likely due to not leaving home much.
- 191 has few day gap but OK
- 203 missing large gap starting on 5/20. Could consider setting study end to 5/19
- 231 has some missing days but OK
- 238 has one four day gap but OK
- 241 has one big gap - Missing: 9/05-9/11/19.   Lapses before, during and after gap.


### General

- All final timezones are in America/Chicago timezone.    

## Set up

## Environment
```{r}
#| message: false
#| warning: false

options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true", 
                      sha1 = "a58e57da996d1b70bb9a5b58241325d6fd78890f")
```

### Paths
```{r}
path_processed <- format_path("studydata/risk/data_processed/shared")
path_ema <- format_path("studydata/risk/data_processed/ema")
path_gps <- format_path("studydata/risk/data_processed/gps")
```

## Packages and plot settings
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)

# adjust plot visuals
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max = Inf)
```

## Cross-reference study visit dates with EMA

### Visit/EMA Dates

Read in visit dates file for all subjects, but filter to only include start date, follow-up 1 date, and end date.
```{r}
dates <- read_csv(here::here(path_processed, "visit_dates.csv"), 
                  show_col_types = FALSE) |> 
  select(subid, study_start = start_study, study_end = end_study, followup_1) |>  
  mutate(study_start = force_tz(study_start, tzone = "America/Chicago"),
         study_end = force_tz(study_end, tzone = "America/Chicago"),
         proj_start = study_start,
         proj_end = study_end) |>  
  glimpse()
```

Filter out subids who did not make it to follow-up 1.
```{r}
dates <- dates |>  
  filter(!is.na(followup_1)) |>  
  select(-followup_1) |>  
  glimpse()
```

Filter out excluded subids for lapse reporting issues.
```{r}
dates <- dates |>  
  filter(!subid %in% c(104, 269, 204)) |> 
  glimpse()
```

Filter out excluded subids for problems with GPS.
```{r}
dates <- dates |>  
  filter(!subid %in% c(2, 21, 51, 56, 190)) |>  
  glimpse()
```

Adjust start start and end dates for missing GPS.
```{r}
dates$study_start[dates$subid == 19] = as_datetime("2017-11-26", tz = "America/Chicago")
dates$study_end[dates$subid == 19] = as_datetime("2018-01-17", tz = "America/Chicago")

dates$study_start[dates$subid == 28] = as_datetime("2017-12-18", tz = "America/Chicago")

dates$study_end[dates$subid == 30] = as_datetime("2017-12-24", tz = "America/Chicago")

dates$study_end[dates$subid == 34] = as_datetime("2018-03-01", tz = "America/Chicago")

dates$study_start[dates$subid == 39] = as_datetime("2018-03-02", tz = "America/Chicago")

dates$study_end[dates$subid == 40] = as_datetime("2018-04-29", tz = "America/Chicago")

dates$study_start[dates$subid == 42] = as_datetime("2018-04-11", tz = "America/Chicago")

dates$study_end[dates$subid == 47] = as_datetime("2018-04-16", tz = "America/Chicago")

dates$study_start[dates$subid == 65] = as_datetime("2018-07-18", tz = "America/Chicago")

dates$study_start[dates$subid == 74] = as_datetime("2018-07-19", tz = "America/Chicago")

dates$study_end[dates$subid == 76] = as_datetime("2018-08-20", tz = "America/Chicago")

dates$study_end[dates$subid == 78] = as_datetime("2018-08-28", tz = "America/Chicago")

dates$study_end[dates$subid == 80] = as_datetime("2018-07-31", tz = "America/Chicago")

dates$study_end[dates$subid == 84] = as_datetime("2018-08-06", tz = "America/Chicago")

dates$study_start[dates$subid == 117] = as_datetime("2018-10-14", tz = "America/Chicago")

dates$study_start[dates$subid == 134] = as_datetime("2018-09-28", tz = "America/Chicago")

dates$study_start[dates$subid == 167] = as_datetime("2019-02-06", tz = "America/Chicago")

dates$study_end[dates$subid == 171] = as_datetime("2019-03-20", tz = "America/Chicago")

dates$study_start[dates$subid == 178] = as_datetime("2019-02-25", tz = "America/Chicago")

dates$study_end[dates$subid == 207] = as_datetime("2019-05-19", tz = "America/Chicago")

dates$study_start[dates$subid == 214] = as_datetime("2019-04-17", tz = "America/Chicago")

dates$study_end[dates$subid == 218] = as_datetime("2019-06-04", tz = "America/Chicago")

dates$study_end[dates$subid == 265] = as_datetime("2019-10-07", tz = "America/Chicago")

dates$study_start[dates$subid == 270] = as_datetime("2019-09-19", tz = "America/Chicago")
```


### Data Start Date

Add column for data_start - will be = study start for EMA b/c everyone started EMA
with study start. Keeping this new column because we use data_start for other studies
and some study generic code may expect it.
```{r}
dates <- dates |>  
  mutate(data_start = study_start) |> 
  glimpse()
```


### EMA end date

Get last EMA data for EMA end.
```{r}
ema <- read_csv(here::here(path_processed, "ema_morning.csv"), show_col_types = FALSE) |>  
    select(subid, start_date)  |>   
  bind_rows(read_csv(here::here(path_processed, "ema_later.csv"), show_col_types = FALSE) |>  
    select(subid, start_date)) |> 
  mutate(subid = as.numeric(subid),
         start_date = with_tz(start_date, tzone = "America/Chicago")) |>  
  rename(ema_end = start_date) |> 
  arrange(subid, desc(ema_end)) |> 
  group_by(subid) |>  
  slice(1) |>  
  glimpse()
```


Add `ema_end` date.
```{r}
dates <- dates |>  
  left_join(ema, by = "subid") |>  
  relocate(subid, study_start, data_start) |>  
  glimpse()
```

### GPS start and end dates

Get first and last GPS date for GPS start and end.
```{r}
gps <- read_csv(here::here(path_processed, "gps.csv"), show_col_types = FALSE) |>  
  group_by(subid) |> 
  arrange(time) |> 
  slice(1, n()) |>
  mutate(start_gps = first(time),
         end_gps = last(time)) |> 
  select(subid, start_gps, end_gps) |> 
  distinct()
```

Add `gps_start` and `gps_end` date.
```{r}
dates <- dates |>  
  left_join(gps, by = "subid") |>  
  relocate(subid, study_start, data_start) |>  
  glimpse()
```

## Save

Save dates.
```{r}
dates  |>  
  write_csv(here::here(path_gps, "study_dates.csv")) |>  
  glimpse()
```

