---
title: "Make GPS demographics"
editor: visual
---

# Housekeeping

## Code Status

In progress as of 07/2024.

## Conclusions

# Set up

## Environment

```{r}
#| message: false
#| warning: false

options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true", 
                      sha1 = "a58e57da996d1b70bb9a5b58241325d6fd78890f")
```

## Paths

```{r}
path_shared <- format_path("studydata/risk/data_processed/shared")
path_gps <- format_path("studydata/risk/data_processed/gps")
path_maps <- format_path("studydata/risk/data_processed/gps/figures")
path_notes <- format_path("GitHub/analysis_risk/shared/notes")
```

## Packages and plot settings

```{r}
#| message: false
#| warning: false

library("tidyverse")
library("scales")
library("ggplot2")

# adjust plot visuals
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max = Inf)
```

# Examine demographic characteristics

## Load in study dates

Load in study_dates.

```{r}
study_dates <- read_csv(file.path(path_gps,"study_dates.csv"),
                        show_col_types = FALSE)
```

Pull list of subject IDs with at least one month data, credible lapse reporting, and sufficient GPS data.

```{r}
subids_dates <- study_dates |>  
  pull(subid) |>  
  unique()
```

## Open and process ID data file containing demographics

```{r}
dem_data <- read_csv(file.path(path_shared, "screen.csv"),
                     show_col_types = FALSE) |> 
  select(subid,  
         age = dem_1, 
         sex = dem_2, 
         race = dem_3, 
         hispanic = dem_4,
         education = dem_5,
         work = dem_6,
         income = dem_7,
         contains("dsm")) |> 
         mutate(hispanic = if_else(str_detect(hispanic, "Yes"), "1", "0")) |>  
    mutate(american_native = ifelse(str_detect(race,"American Indian/Alaska Native"),1,0),
        asian = ifelse(str_detect(race,"Asian"),1,0),
        pacific = ifelse(str_detect(race,"Native Hawaiian or Other Pacific Islander"),1,0),
        black = ifelse(str_detect(race,"Black/African American"),1,0),
        white = ifelse(str_detect(race,"White/Caucasian"),1,0),
        other = ifelse(str_detect(race,"Other/Multiracial"),1,0),
        )
```

Filter down demographics file to only include subjects being used in GPS analyses.

```{r}
dem_data <-  dem_data |>  
  filter(subid %in% subids_dates)
```

## Summary tables

### Age

```{r}
age_summary <- dem_data |> 
  summarize(mean=round(mean(age),1), sd=round(sd(age),1), min=min(age), max=max(age))
age_summary
```

### Sex

```{r}
sex_summary <- dem_data |> 
  group_by(sex) |> 
  summarize(n = n()) |> 
  mutate(`%` = scales::percent(n/sum(n)))
sex_summary
```

### Race

```{r}
race_summary <- dem_data |> 
  group_by(race) |> 
  summarize(n = n()) |> 
  mutate(`%` = scales::percent(n/sum(n)))
race_summary
```

### Ethnicity

```{r}
hisp_summary <- dem_data |> 
  group_by(hispanic) |> 
  summarize(n = n()) |> 
  mutate(`%` = scales::percent(n/sum(n)))
hisp_summary  
```

### Income

```{r}
income_summary <- dem_data |>
  drop_na() |> # probably need a better way to do this
  summarize(mean=scales::dollar(round(mean(income))), sd=scales::dollar(round(sd(income))), min=scales::dollar(min(income)), max=scales::dollar(max(income)))
income_summary
```

### DSM

Total DSM score

```{r}
dsm_summary <- dem_data |>  
  mutate(across(dsm5_1:dsm5_11, ~ recode(., "No" = 0, "Yes" = 1))) |>  
  rowwise() |>  
  mutate(dsm5_total = sum(c(dsm5_1, dsm5_2, dsm5_3, dsm5_4, dsm5_5, dsm5_6, dsm5_7, 
                            dsm5_8, dsm5_9, dsm5_10, dsm5_11))) |>  
  ungroup() |>  
  summarise(min = min(dsm5_total),
            max = max(dsm5_total),
            mean = mean(dsm5_total))
dsm_summary
```

## Figures

### Age

```{r}
dem_data |> 
  ggplot(aes(x = age)) +
  geom_density(fill = "blue", color = "darkblue", alpha = .1, size = .8)
```

### Sex

```{r}
dem_data |> 
  group_by(sex) |> 
  summarize(n = n()) |>
  ggplot(aes(x="", y=n, fill=sex)) +
  geom_bar(stat="identity", color = "black", size = 1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values = c("white", "gray")) +
  theme_void()
```

### Race

```{r}
dem_data |> 
  group_by(race) |> 
  summarize(n = n()) |>
  ggplot(aes(x="", y=n, fill=race)) +
  geom_bar(stat="identity", color = "black", size = 1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values = c("#3a3a3a", "#838383", 
                               "#aeaeae", "#c9c9c9",
                               "#e5e5e5", "white")) +
  theme_void()
```

### Ethnicity

```{r}
dem_data |>
  mutate(hispanic = recode(hispanic,
                           `0` = "No",
                           `1` = "Yes")) |> 
  group_by(hispanic) |> 
  summarize(n = n()) |>
  ggplot(aes(x="", y=n, fill=hispanic)) +
  geom_bar(stat="identity", color = "black", size = 1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values = c("white", "gray")) +
  theme_void()
```

### Income

```{r}
dem_data |> 
  ggplot(aes(x = income)) +
  geom_density(fill = "blue", color = "darkblue", alpha = .1, size = .8)
```

### DSM

```{r}
dem_data |>
  mutate(across(dsm5_1:dsm5_11, ~ recode(., "No" = 0, "Yes" = 1))) |>  
  rowwise() |>  
  mutate(dsm5_total = sum(c(dsm5_1, dsm5_2, dsm5_3, dsm5_4, dsm5_5, dsm5_6, dsm5_7, 
                            dsm5_8, dsm5_9, dsm5_10, dsm5_11))) |> 
  ungroup() |> 
  group_by(dsm5_total) |> 
  summarize(n = n()) |>
  ggplot(aes(x = dsm5_total, y = n)) +
  geom_bar(stat = "identity", color = "black", fill = "white")
```
