---
title: "Make auROCs for Model Fairness Evaluations"
author: "Updated by CP from Kendra Wyant for GPS"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  version: "v25"
  cv: "nested_2_x_5_6_x_5"
  model: "full_v_baseline"
  batch: "full"
editor_options: 
  chunk_output_type: console
---

Notes:

This is theoretically set up to work but just need to think about how to add in labels.

### Set Up Environment

Packages, functions, and paths
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_shared <- format_path(str_c("risk/data_processed/shared"))
path_models <- format_path(str_c("risk/models/gps"))

version <- params$version
cv <- params$cv
model <- params$model
batch <- params$batch
```


### Load model predictions

```{r}
preds <- read_csv(here::here(path_models, str_c("outer_preds_", 
                                            version, "_", 
                                            cv, "_", batch, ".csv")))


#preds <- read_rds(here::here(path_models, "outer_preds_6_x_5_1_x_5_day_v10_nested_full.rds"))
#preds_final <- read_rds(here::here(path_models, "preds_1_x_5_day_v10_kfold_full_final.rds"))
```

### Load fairness features

```{r}
labels <- read_csv(here::here(path_shared, "labels_gps_day_24h.csv"),
                   show_col_types = FALSE) |> 
  mutate(label_num = 1:n()) # CP think more about this, not sure if best place to add 

dem <- read_csv(here::here(path_shared, "features_demographics.csv"),
                 show_col_types = FALSE) |> 
  filter(subid %in% labels$subid) |> 
  mutate(demo_educ_comb = if_else(demo_educ == "College or more",
                                  "College or more",
                                  "Less than college"),
         income_split = if_else(demo_income > 25100,
                                "Above 2018 Poverty Line",
                                "Below 2018 Povery Line"),
         age_split = if_else(demo_age > 55, "Older than 55", "Younger than 55"))

#dem <- read_csv(here::here(path_shared, "features_fairness.csv"),
                 #show_col_types = FALSE) |>
  #select(-ruca1) |> 
  #filter(subid %in% labels$subid)
```

### Link features with predictions

```{r}
feat_preds <- labels |> 
  select(id_obs = label_num,
         label = lapse,
         subid) |>
  right_join(preds, by = c("id_obs", "label")) |> 
  mutate(label = factor(label, levels = c("lapse", "no lapse"))) |> 
  left_join(dem, by = "subid")

# nrow(labels)
# nrow(preds) # should be nrow(labels) * 6 (number of outer repeats)
# nrow(feat_preds) # should be same as nrow(preds)

# feat_preds_final <- labels |> 
#   select(id_obs = label_num,
#          label = lapse,
#          subid) |>
#   right_join(preds_final, by = c("id_obs", "label")) |> 
#   mutate(label = factor(label, levels = c("Lapse", "No lapse"))) |> 
#   left_join(dem, by = "subid")
# 
# # nrow(labels)
# # nrow(preds_final) # should be same as nrow(labels)
# # nrow(feat_preds_final) # should be same as nrow(preds_final)
# 
# feat_preds_final |> 
#   write_rds(here::here(path_models, "fairness_preds_1_x_5_day_v10_kfold_full_final.rds"))
```


### Calculate auROCs by group

This calculates auROC using test sets and raw probabilities.   

function
```{r}
get_auroc_by_dem <- function(df, var) {
  # df should be dataframe with demographic features and predictions
  # var should be character or factor variable in df
  
  var_tmp <- df |> 
      pull(get(var))
   
  var_levels <- 
     if (class(var_tmp) != "factor") {
       levels(as.factor(var_tmp))
     } else levels(var_tmp)
  
  # create empty tibble with split nums
  auroc <- tibble(outer_split_num = unique(df$outer_split_num)) |> 
    arrange(outer_split_num) 
    
  for (level in var_levels) {
     auroc_level <- df |>
       filter(get(var) == level) |> 
       nest(.by = outer_split_num, .key = "preds") |> 
       mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
       select(-preds) |> 
       unnest(auroc) |> 
       select(-c(.estimator, .metric)) |> 
       rename(!!level:=.estimate)
     
     auroc <- auroc |> 
       left_join(auroc_level, by = "outer_split_num")
  }
  
  return(auroc)
}
```


get auROCs 
```{r}
auroc_dem <- get_auroc_by_dem(df = feat_preds, var = "demo_sex") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "demo_race"), by = "outer_split_num") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "income_split"), by = "outer_split_num") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "age_split"), by = "outer_split_num")  |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "demo_educ_comb"), by = "outer_split_num")  
```

> CP note: is it helpful or useful to get LL by demographic group? my understanding is that LL is calculated for each observation, so I'm not sure if this is 

```{r}
get_ll_by_dem <- function(df, var) {
  # df should be dataframe with demographic features and predictions
  # var should be character or factor variable in df
  
  var_tmp <- df |> 
      pull(get(var))
   
  var_levels <- 
     if (class(var_tmp) != "factor") {
       levels(as.factor(var_tmp))
     } else levels(var_tmp)
  
  # create empty tibble with split nums
  ll <- tibble(outer_split_num = unique(df$outer_split_num)) |> 
    arrange(outer_split_num) 
    
  for (level in var_levels) {
     ll_level <- df |>
       filter(get(var) == level) |> 
       nest(.by = outer_split_num, .key = "preds") |> 
       mutate(ll = map(preds, \(preds) mn_log_loss(preds, prob_raw, 
                                             truth = label))) |> 
       select(-preds) |> 
       unnest(ll) |> 
       select(-c(.estimator, .metric)) |> 
       rename(!!level:=.estimate)
     
     ll <- ll |> 
       left_join(ll_level, by = "outer_split_num")
  }
  
  return(ll)
}
```

get LLs
```{r}
ll_dem <- get_ll_by_dem(df = feat_preds, var = "demo_sex") |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "demo_race"), by = "outer_split_num") |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "income_split"), by = "outer_split_num") |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "age_split"), by = "outer_split_num")  |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "demo_educ_comb"), by = "outer_split_num")  
```

Save auroc_dem
```{r}
#auroc_dem |> 
#  write_csv(here::here(path_models, "fairness_auroc_6_x_5_1_x_5_nested_v10_full.csv"))  
```
