---
title: "Model probabilities"
author: "John Curtin, edited by Claire for study_gps"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  version: "v24"
  cv: "nested_2_x_5_6_x_5"
  model: "full_v_baseline"
  batch: "full"
editor_options: 
  chunk_output_type: console
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

options(conflicts.policy = "depends.ok")
library(probably)

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_input <- format_path("risk/models/gps")

version <- params$version
cv <- params$cv
model <- params$model
batch <- params$batch
```

```{r}
preds <- read_csv(here::here(path_input, str_c("outer_preds_", 
                                            version, "_", 
                                            cv, "_", batch, ".csv")))

metrics <- read_csv(here::here(path_input, str_c("outer_metrics_", 
                                           version, "_", 
                                           cv, "_", batch, ".csv")))

```

## Rank order and compare raw and platt calibration

Smoothed
```{r}
cor(preds$prob_raw, preds$prob_logi_smooth, method = "spearman")

plot(preds$prob_raw, qlogis(preds$prob_logi_smooth))

preds |>
  group_by(outer_split_num) |>
  summarise(
    spear = cor(prob_raw, prob_logi_smooth, method = "spearman")
  ) |>
  arrange(spear)
```

Not smoothed
 - check warning

```{r}
cor(preds$prob_raw, preds$prob_logi_no_smooth, method = "spearman")

plot(preds$prob_raw, qlogis(preds$prob_logi_no_smooth))

preds |>
  group_by(outer_split_num) |>
  summarise(
    spear = cor(prob_raw, prob_logi_no_smooth, method = "spearman")
  ) |>
  arrange(spear)
```

Correlate smoothed and unsmoothed.
```{r}
cor(preds$prob_logi_smooth, preds$prob_logi_no_smooth)

cor(preds$prob_logi_smooth, preds$prob_raw)
```

## Basic histogram and calibration plot

Get lapse prevalence and log loss in each fold.
```{r}
text_labels <- preds |> 
  group_by(outer_split_num) |> 
  mutate(prev = mean(label == "lapse"),
         log_loss_raw = mn_log_loss_vec(truth = factor(label,
                                                   levels = c("lapse",
                                                              "no lapse")),
                                    estimate = prob_raw)) |> 
  ungroup() |> 
  merge(metrics, by = "outer_split_num") |> 
  select(outer_split_num, prev, log_loss_raw, hp1, hp2, hp3,
         resample, roc_auc) |> 
  distinct(outer_split_num, .keep_all = TRUE)

text_labels <- text_labels |>
  mutate(
    label = paste0(
      "prevalence: ", round(prev, 3),
      "\nlog loss: ", round(log_loss_raw, 3),
      "\nhp1: ", hp1,
      "\nhp2: ", hp2,
      "\nhp3: ", hp3,
      "\nresample: ", resample,
      "\nauroc: ", round(roc_auc, 3)
    )
  )
```
 
Create faceted histograms.
```{r}
preds |> 
  ggplot(aes(x = prob_raw)) +
    geom_histogram(binwidth = .01) +
    labs(title = "Histogram of Raw Probabilities") + 
    scale_x_continuous(breaks = seq(0, 1, .25)) +
    coord_cartesian(xlim = c(0, 1)) +
    facet_wrap(~ outer_split_num) +
  geom_text(
  data = text_labels,
  aes(x = Inf, y = Inf, label = label),
  hjust = 1.05, vjust = 1.1, size = 2.5) +
  theme_light()

preds |> 
  ggplot(aes(x = prob_logi_smooth)) +
    geom_histogram(binwidth = .01) +
    labs(title = "Histogram of Logi Probabilities (Smoothed)") + 
    scale_x_continuous(breaks = seq(0, 1, .25)) +
    coord_cartesian(xlim = c(0, 1)) +
    facet_wrap(~ outer_split_num) +
    geom_text(
    data = text_labels,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.05, vjust = 1.1, size = 2.5) +
    theme_light()

preds |> 
  ggplot(aes(x = prob_logi_no_smooth)) +
    geom_histogram(binwidth = .01) +
    labs(title = "Histogram of Logi Probabilities (Not Smoothed)") + 
    scale_x_continuous(breaks = seq(0, 1, .25)) +
    coord_cartesian(xlim = c(0, 1)) +
    facet_wrap(~ outer_split_num) +
    geom_text(
    data = text_labels,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.05, vjust = 1.1, size = 2.5) +
    theme_light()
```

Calibration plot (original)
```{r}
bin_width = 0.10
cal_raw <- preds |>
  mutate(
    prob = prob_raw,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "raw"
  )

cal_logi_smooth <- preds |>
  mutate(
    prob = prob_logi_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_smooth"
  )

cal_logi_no_smooth <- preds |>
  mutate(
    prob = prob_logi_no_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_no_smooth"
  )


cal_raw |> 
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |> 
  mutate(model = fct(probs, levels = c("raw", "logi_smooth", "logi_no_smooth"))) |> 
  group_by(bins, probs)  |>
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse, color = probs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_line() +
    geom_point() +
    labs(x = "Predicted Lapse Probability (bin mid-point)",
         y = "Observed Lapse Probability",
         title = "Calibration (Raw v Platt, Smoothed and Unsmoothed)") +
    scale_x_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1))

```

Create a plot of probabilities per fold, so we can see the variability within folds.
```{r}
bin_width = 0.10
cal_raw <- preds |>
  mutate(
    prob = prob_raw,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "raw"
  )

cal_logi_smooth <- preds |>
  mutate(
    prob = prob_logi_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_smooth"
  )

cal_logi_no_smooth <- preds |>
  mutate(
    prob = prob_logi_no_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_no_smooth"
  )


cal_raw |> 
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |> 
  mutate(model = fct(probs, levels = c("raw", "logi_smooth", "logi_no_smooth"))) |> 
  group_by(outer_split_num, bins, probs)  |>
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse, color = probs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_line() +
    geom_point() +
    labs(x = "Predicted Lapse Probability (bin mid-point)",
         y = "Observed Lapse Probability",
         title = "Calibration (Raw v Platt, Smoothed and Unsmoothed)") +
    scale_x_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
  facet_wrap(~ outer_split_num)

```

Check how many observations are in a given bin -- this might help account for the instability we see across folds. This is collapsing across folds...
```{r}
cal_raw |>
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |>
  group_by(bins, probs) |>
  summarise(
    mean_pred = mean(prob),
    mean_lapse = mean(lapse),
    n = n(),
    .groups = "drop"
  )
```

...but we can also look at this grouped by folds to get better insight into our plots. It looks like part of the issue with looking at our calibration plots is that some of our bins have very few observations.
```{r}
cal_raw |>
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |>
  group_by(bins, probs, outer_split_num) |>
  summarise(
    mean_pred = mean(prob),
    mean_lapse = mean(lapse),
    n = n(),
    .groups = "drop"
  ) |> 
  view()
```

We can also print a table of the absolute difference between predicted and true values of lapse for each probability type. Platt scaling without smoothing seems to produce the lowest error.
```{r}
cal_raw |>
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |>
  group_by(outer_split_num, probs, bins) |>
  summarize(
    mean_lapse = mean(lapse),
    mean_pred = mean(prob),
    .groups = "drop"
  ) |>
  mutate(
    calibration_error = abs(mean_pred - mean_lapse)
  ) |>
  group_by(probs) |>
  summarize(
    mean_cal_error = mean(calibration_error),
    median_cal_error = median(calibration_error),
    max_cal_error = max(calibration_error)
  )
```

## Brier scores

```{r}
brier_scores <- preds |>
  pivot_longer(
    cols = c(prob_raw, prob_logi_smooth, prob_logi_no_smooth),
    names_to = "method",
    values_to = ".pred_lapse") |>
  mutate(label = factor(label,
                        levels = c("lapse", "no lapse"))) |> 
  group_by(outer_split_num,
           method) |>
  brier_class(label, .pred_lapse)
```

Observations:
 - Platt without smoothing is the closest to observed probabilities
 - It also has the smallest SD (stable across folds)
```{r}
brier_scores |>
  group_by(method) |>
  summarise(
    mean_brier = mean(.estimate),
    sd_brier = sd(.estimate),
    median_brier = median(.estimate),
    min_brier = min(.estimate),
    max_brier = max(.estimate)
  )
```

## KW Calibration

Create histogram comparing raw/platt across lapse/no lapse cases.
```{r}
preds |> 
  pivot_longer(cols = c(prob_raw, prob_logi), names_to = "prob_type", values_to = "prob") |>
  mutate(prob_type = factor(prob_type, levels = c("prob_raw", "prob_logi"),
                                labels = c("Raw (uncalibrated)", "Platt calibration"))) |> 
  ggplot() +
  geom_histogram(aes(x = prob, fill = prob_type), color = "white", bins = 40) +
  facet_grid(label ~ prob_type, scales = "free_y") +
  labs(x = "Probabilties",
       y = NULL) +
  scale_fill_manual(values = c("#240e31", "#046B52")) +
  theme_classic() +
  theme(axis.text=element_text(size=9.5),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "none")
```

Set up data to add rug to plot -- currently doesn't display properly on plot.
```{r}
bin_width = 0.10

preds_all_logi <- preds |> 
  mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "lapse", 1, 0),
         prob = "Platt calibration")  

preds_all_raw <- preds |> 
  mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "lapse", 1, 0),
         prob = "Raw (uncalibrated)") 

rug_data <- preds_all_raw |> 
  mutate(
    bins = as.numeric(bins),
    rug_center = bin_width/2 + bin_width * (bins - 1),
    rug_x = rug_center + runif(n(), -bin_width/2, bin_width/2)
  )
```

Create calibration plot.
```{r}
preds_all_raw |> 
  bind_rows(preds_all_logi) |> 
  mutate(prob = factor(prob, levels = c("Raw (uncalibrated)", 
                                        "Platt calibration"))) |> 
  group_by(bins,  prob)  |> 
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |> 
  mutate(
    bins = as.numeric(bins),
    midpoints = bin_width/2 + bin_width * (bins - 1)
  )  |> 
  ggplot(aes(x = midpoints, y = mean_lapse, group = prob,
             color = prob, linetype = prob)) +
  geom_abline(slope = 1, intercept = 0, linetype = "longdash", color = "gray80") +
  geom_line(linewidth = .7) +
  labs(
    x = "Predicted Lapse Probability (Bin Midpoint)",
    y = "Observed Lapse Probability",
    color = NULL,
    linetype = NULL
  ) +
  ## currently makes this run super slow, think bc it's very condensed
  #geom_rug(data = rug_data, aes(x = rug_x), sides = "b", alpha = 0.3,
           #inherit.aes = FALSE) +
  scale_x_continuous(breaks = seq(0, .8, bin_width), limits = c(0, .8)) +
  scale_y_continuous(limits = c(-0.08, .8), breaks = seq(0,.8, bin_width),
                     expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  scale_color_manual(values = c("#240e31", "#046B52")) +
  scale_linetype_manual(values = c("solid", "twodash")) +
  theme_classic() +
  theme(axis.text=element_text(size=9.5),
        legend.key.size = unit(0.25, "cm"),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "bottom") 
```

