---
title: "Model probabilities"
author: "John Curtin, edited by Claire for study_gps"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  version: "v24"
  cv: "nested_2_x_5_6_x_5"
  model: "full_v_baseline"
  batch: "full"
editor_options: 
  chunk_output_type: console
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

options(conflicts.policy = "depends.ok")
library(probably)

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_input <- format_path("risk/models/gps")

version <- params$version
cv <- params$cv
model <- params$model
batch <- params$batch
```

Brier score function.
```{r}
get_brier <- function(split, df){
  
  #filter to split
  probs_split <- df |> 
    filter(outer_split_num == split)

  raw <- probs_split |>
    filter(method == "prob_raw") |> 
    brier_class(label, .pred_lapse)
  
  logi_smooth <- probs_split |>
    filter(method == "prob_logi_smooth") |> 
    brier_class(label, .pred_lapse)
  
  logi_no_smooth <- probs_split |>
    filter(method == "prob_logi_no_smooth") |> 
    brier_class(label, .pred_lapse)
  
  #iso <- probs_split |>
    #filter(method == "prob_iso") |> 
    #brier_class(label, .pred_Lapse)
  
  #beta <- probs_split |>
    #filter(method == "prob_beta") |> 
    #brier_class(label, .pred_Lapse)
  
  brier <- tibble(raw = raw$.estimate, 
                 logi_smooth = logi_smooth$.estimate,
                 logi_no_smooth = logi_no_smooth$.estimate) 
                 #iso = iso$.estimate) 
  #               beta = beta$.estimate)
  
  #brier <- tibble(raw = raw$.estimate)
  
  return(brier)
}
```

```{r}
preds <- read_csv(here::here(path_input, str_c("outer_preds_", 
                                            version, "_", 
                                            cv, "_", batch, ".csv")))

```

## Rank order and compare raw and platt calibration

Smoothed
```{r}
cor(preds$prob_raw, preds$prob_logi_smooth, method = "spearman")

plot(preds$prob_raw, qlogis(preds$prob_logi_smooth))

preds |>
  group_by(outer_split_num) |>
  summarise(
    spear = cor(prob_raw, prob_logi_smooth, method = "spearman")
  ) |>
  arrange(spear)
```

Not smoothed
```{r}
cor(preds$prob_raw, preds$prob_logi_no_smooth, method = "spearman")

plot(preds$prob_raw, qlogis(preds$prob_logi_no_smooth))

preds |>
  group_by(outer_split_num) |>
  summarise(
    spear = cor(prob_raw, prob_logi_no_smooth, method = "spearman")
  ) |>
  arrange(spear)
```

## Basic histogram and calibration plot

Histogram of raw probabilities
```{r}
preds |> 
  ggplot(aes(x = prob_raw)) +
    geom_histogram(binwidth = .01) +
    labs(title = "Histogram of Raw Probabilities") + 
    scale_x_continuous(breaks = seq(0, 1, .25), limits = c(0, 1))

preds |> 
  ggplot(aes(x = prob_logi_smooth)) +
    geom_histogram(binwidth = .01) +
    labs(title = "Histogram of Logi Probabilities (Smoothed)") + 
    scale_x_continuous(breaks = seq(0, 1, .25), limits = c(0, 1))

preds |> 
  ggplot(aes(x = prob_logi_no_smooth)) +
    geom_histogram(binwidth = .01) +
    labs(title = "Histogram of Logi Probabilities (Not Smoothed)") + 
    scale_x_continuous(breaks = seq(0, 1, .25), limits = c(0, 1))
```

Calibration plot (original)
```{r}
bin_width = 0.10
cal_raw <- preds |>
  mutate(
    prob = prob_raw,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "raw"
  )

cal_logi_smooth <- preds |>
  mutate(
    prob = prob_logi_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_smooth"
  )

cal_logi_no_smooth <- preds |>
  mutate(
    prob = prob_logi_no_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_no_smooth"
  )


cal_raw |> 
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |> 
  mutate(model = fct(probs, levels = c("raw", "logi_smooth", "logi_no_smooth"))) |> 
  group_by(bins, probs)  |>
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse, color = probs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_line() +
    geom_point() +
    labs(x = "Predicted Lapse Probability (bin mid-point)",
         y = "Observed Lapse Probability",
         title = "Calibration (Raw v Platt, Smoothed and Unsmoothed)") +
    scale_x_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1))

```

```{r}
cal_tbl <- cal_raw |>
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |>
  group_by(bins, probs) |>
  summarise(
    mean_pred = mean(prob),
    mean_lapse = mean(lapse),
    n = n(),
    .groups = "drop"
  )

cal_tbl
```

```{r}
bin_width = 0.10
cal_raw <- preds |>
  mutate(
    prob = prob_raw,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "raw"
  )

cal_logi_smooth <- preds |>
  mutate(
    prob = prob_logi_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_smooth"
  )

cal_logi_no_smooth <- preds |>
  mutate(
    prob = prob_logi_no_smooth,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "logi_no_smooth"
  )


cal_raw |> 
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |> 
  mutate(model = fct(probs, levels = c("raw", "logi_smooth", "logi_no_smooth"))) |> 
  group_by(outer_split_num, bins, probs)  |>
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse, color = probs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_line() +
    geom_point() +
    labs(x = "Predicted Lapse Probability (bin mid-point)",
         y = "Observed Lapse Probability",
         title = "Calibration (Raw v Platt, Smoothed and Unsmoothed)") +
    scale_x_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
  facet_wrap(~ outer_split_num)

```

```{r}
cal_error <- cal_raw |>
  bind_rows(cal_logi_smooth, cal_logi_no_smooth) |>
  group_by(outer_split_num, probs, bins) |>
  summarize(
    mean_lapse = mean(lapse),
    mean_pred = mean(prob),
    .groups = "drop"
  ) |>
  mutate(
    calibration_error = abs(mean_pred - mean_lapse)
  )

cal_error |>
  group_by(probs) |>
  summarize(
    mean_cal_error = mean(calibration_error),
    median_cal_error = median(calibration_error),
    max_cal_error = max(calibration_error)
  )
```

## Brier scores

```{r}
brier_scores <- preds |>
  pivot_longer(
    cols = c(prob_raw, prob_logi_smooth, prob_logi_no_smooth),
    names_to = "method",
    values_to = ".pred_lapse"
  ) |>
  mutate(label = factor(label, levels = c("lapse", "no lapse"))) |> 
  group_by(
    outer_split_num,
    method
  ) |>
  brier_class(
    label,
    .pred_lapse
  )
```

```{r}
brier_scores |>
  group_by(method) |>
  summarise(
    mean_brier = mean(.estimate),
    sd_brier = sd(.estimate),
    median_brier = median(.estimate),
    min_brier = min(.estimate),
    max_brier = max(.estimate)
  )

```

## KW Calibration

Create histogram comparing raw/platt across lapse/no lapse cases.
```{r}
preds |> 
  pivot_longer(cols = c(prob_raw, prob_logi), names_to = "prob_type", values_to = "prob") |>
  mutate(prob_type = factor(prob_type, levels = c("prob_raw", "prob_logi"),
                                labels = c("Raw (uncalibrated)", "Platt calibration"))) |> 
  ggplot() +
  geom_histogram(aes(x = prob, fill = prob_type), color = "white", bins = 40) +
  facet_grid(label ~ prob_type, scales = "free_y") +
  labs(x = "Probabilties",
       y = NULL) +
  scale_fill_manual(values = c("#240e31", "#046B52")) +
  theme_classic() +
  theme(axis.text=element_text(size=9.5),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "none")
```

Set up data to add rug to plot -- currently doesn't display properly on plot.
```{r}
bin_width = 0.10

preds_all_logi <- preds |> 
  mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "lapse", 1, 0),
         prob = "Platt calibration")  

preds_all_raw <- preds |> 
  mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "lapse", 1, 0),
         prob = "Raw (uncalibrated)") 

rug_data <- preds_all_raw |> 
  mutate(
    bins = as.numeric(bins),
    rug_center = bin_width/2 + bin_width * (bins - 1),
    rug_x = rug_center + runif(n(), -bin_width/2, bin_width/2)
  )
```

Create calibration plot.
```{r}
preds_all_raw |> 
  bind_rows(preds_all_logi) |> 
  mutate(prob = factor(prob, levels = c("Raw (uncalibrated)", 
                                        "Platt calibration"))) |> 
  group_by(bins,  prob)  |> 
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |> 
  mutate(
    bins = as.numeric(bins),
    midpoints = bin_width/2 + bin_width * (bins - 1)
  )  |> 
  ggplot(aes(x = midpoints, y = mean_lapse, group = prob,
             color = prob, linetype = prob)) +
  geom_abline(slope = 1, intercept = 0, linetype = "longdash", color = "gray80") +
  geom_line(linewidth = .7) +
  labs(
    x = "Predicted Lapse Probability (Bin Midpoint)",
    y = "Observed Lapse Probability",
    color = NULL,
    linetype = NULL
  ) +
  ## currently makes this run super slow, think bc it's very condensed
  #geom_rug(data = rug_data, aes(x = rug_x), sides = "b", alpha = 0.3,
           #inherit.aes = FALSE) +
  scale_x_continuous(breaks = seq(0, .8, bin_width), limits = c(0, .8)) +
  scale_y_continuous(limits = c(-0.08, .8), breaks = seq(0,.8, bin_width),
                     expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  scale_color_manual(values = c("#240e31", "#046B52")) +
  scale_linetype_manual(values = c("solid", "twodash")) +
  theme_classic() +
  theme(axis.text=element_text(size=9.5),
        legend.key.size = unit(0.25, "cm"),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "bottom") 
```

