---
title: "Make auROCs for Model Fairness Evaluations"
author: "Updated by CP from Kendra Wyant for GPS"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  version: "v25"
  cv: "nested_2_x_5_6_x_5"
  model: "full_v_baseline"
  batch: "full"
editor_options: 
  chunk_output_type: console
---

### Set Up Environment

Packages, functions, and paths
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
library(tidyposterior)

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_shared <- format_path(str_c("risk/data_processed/shared"))
path_models <- format_path(str_c("risk/models/gps"))

version <- params$version
cv <- params$cv
model <- params$model
batch <- params$batch
```


### Load model predictions

```{r}
preds <- read_csv(here::here(path_models, str_c("outer_preds_", 
                                            version, "_", 
                                            cv, "_", batch, ".csv")))


#preds <- read_rds(here::here(path_models, "outer_preds_6_x_5_1_x_5_day_v10_nested_full.rds"))
#preds_final <- read_rds(here::here(path_models, "preds_1_x_5_day_v10_kfold_full_final.rds"))
```

### Load fairness features

```{r}
labels <- read_csv(here::here(path_shared, "labels_gps_day_24h.csv"),
                   show_col_types = FALSE) |> 
  mutate(label_num = 1:n())

dem <- read_csv(here::here(path_shared, "features_demographics.csv"),
                 show_col_types = FALSE) |> 
  filter(subid %in% labels$subid) |> 
  mutate(demo_educ_comb = if_else(demo_educ == "College or more",
                                  "College or more",
                                  "Less than college"),
         income_split = if_else(demo_income > 25100,
                                "Above 2018 Poverty Line",
                                "Below 2018 Poverty Line"),
         age_split = if_else(demo_age > 55, "Older than 55", "Younger than 55"),
         demo_race = if_else(demo_race == "Other", "Non-White and/or Hispanic", demo_race))

#dem <- read_csv(here::here(path_shared, "features_fairness.csv"),
                 #show_col_types = FALSE) |>
  #select(-ruca1) |> 
  #filter(subid %in% labels$subid)
```

### Link features with predictions

> `label_num` was not included in original modeling and so was reproduced here after the fact. This was done through several processes. First, CP reviewed scripts from other projects (e.g., Risk 2, Lag) to determine that each `label_num` corresponds to one individual lapse observation -- what is most important is that the label in the labels file matches the label in the data file. A label identifier is added to the labels dataframe in the code below, which is organized sequentially by `subid` and `dttm_label`. A label identifier was added to the features in code chunk 8 in `2_eval_outer.qmd` following an `arrange()` based on `subid` and `dttm_label`. Thus these are believed to now match each other!

```{r}
feat_preds <- labels |> 
  select(id_obs = label_num,
         label = lapse,
         subid) |>
  right_join(preds, by = c("id_obs", "label")) |> 
  mutate(label = factor(label, levels = c("lapse", "no lapse"))) |> 
  left_join(dem, by = "subid")

# nrow(labels)
# nrow(preds) # should be nrow(labels) * 6 (number of outer repeats)
# nrow(feat_preds) # should be same as nrow(preds)
```


### Calculate auROCs by group

This calculates auROC using test sets and raw probabilities.   

function
```{r}
get_auroc_by_dem <- function(df, var) {
  # df should be dataframe with demographic features and predictions
  # var should be character or factor variable in df
  
  var_tmp <- df |> 
      pull(get(var))
   
  var_levels <- 
     if (class(var_tmp) != "factor") {
       levels(as.factor(var_tmp))
     } else levels(var_tmp)
  
  # create empty tibble with split nums
  auroc <- tibble(outer_split_num = unique(df$outer_split_num)) |> 
    arrange(outer_split_num) 
    
  for (level in var_levels) {
     auroc_level <- df |>
       filter(get(var) == level) |> 
       nest(.by = outer_split_num, .key = "preds") |> 
       mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
       select(-preds) |> 
       unnest(auroc) |> 
       select(-c(.estimator, .metric)) |> 
       rename(!!level:=.estimate)
     
     auroc <- auroc |> 
       left_join(auroc_level, by = "outer_split_num")
  }
  
  return(auroc)
}
```

get auROCs 
```{r}
auroc_dem <- get_auroc_by_dem(df = feat_preds, var = "demo_sex") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "demo_race"), by = "outer_split_num") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "income_split"), by = "outer_split_num") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "age_split"), by = "outer_split_num")  |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "demo_educ_comb"), by = "outer_split_num") |> 
    mutate(fold_num = rep(1:5, 6),
         repeat_num = c(rep(1, 5), rep(2, 5), rep(3, 5), 
                        rep(4, 5), rep(5, 5), rep(6, 5))) |>
  select(-outer_split_num)
```

function
```{r}
calc_pp <- function (data, dem_var) {
  
  data <- 
    if (dem_var == "sex") {
    data |> 
    select(id = repeat_num, id2 = fold_num, Male, Female)
  } else if (dem_var == "race") {
    data |> 
    select(id = repeat_num, id2 = fold_num, `White/Caucasian`, `Non-White and/or Hispanic`)
  } else if (dem_var == "income") {
    data |> 
    select(id = repeat_num, id2 = fold_num, `Above 2018 Poverty Line`, `Below 2018 Poverty Line`)
  } else if (dem_var == "age") {
    data |> 
    select(id = repeat_num, id2 = fold_num, `Younger than 55`, `Older than 55`)
  } else if (dem_var == "education") {
    data |> 
    select(id = repeat_num, id2 = fold_num, `College or more`, `Less than college`)
  } else {
    stop(dem_var, " not in data")
  }
  
  
  set.seed(101)
  pp <- data |> 
    perf_mod(formula = statistic ~ model + (1 | id2/id),
             transform = tidyposterior::logit_trans,  
             iter = 6000, chains = 4,  
             adapt_delta = .999,
             family = gaussian) 

  pp_tidy <- pp |> 
    tidy(seed = 123) 

  q = c(.025, .5, .975)
  ci <- pp_tidy |> 
    group_by(model) |> 
    summarize(pp_median = quantile(posterior, probs = q[2]),
              pp_lower = quantile(posterior, probs = q[1]), 
              pp_upper = quantile(posterior, probs = q[3]))  |> 
    arrange(model)
  
  
  contrast_lists <- 
    if (dem_var == "sex") {
    c(list("Male"), list("Female"))
  } else if (dem_var == "race") {
    c(list("White/Caucasian"), list("Non-White and/or Hispanic"))
  } else if (dem_var == "income") {
    c(list("Above 2018 Poverty Line"), list("Below 2018 Poverty Line"))
  } else if (dem_var == "age") {
    c(list("Younger than 55"), list("Older than 55"))
  } else if (dem_var == "education") {
    c(list("College or more"), list("Less than college"))
  } else {
    stop(dem_var, " not in data")
  }
      
  ci_contrast <- pp |>
    contrast_models(contrast_lists[1],  contrast_lists[2]) |> 
    summary(size = 0) |> 
    mutate(probability = 1 - probability)
  
  ci_median_contrast <- pp |> 
    contrast_models(contrast_lists[1],  contrast_lists[2]) |>  
    group_by(contrast) |> 
    summarize(median = quantile(difference, .5)) |> 
    mutate(contrast = str_remove(contrast, "\\."))


ci_contrast <- ci_contrast |> 
    left_join(ci_median_contrast, by = c("contrast")) |> 
    select(contrast, probability, median, lower, upper) 
  
  list(pp = pp_tidy, ci = ci, ci_contrast = ci_contrast)
}
```

Generate comparisons.
```{r}
sex <- calc_pp(auroc_dem, "sex") 
race <- calc_pp(auroc_dem, "race")
income <- calc_pp(auroc_dem, "income")
age <- calc_pp(auroc_dem, "age")
education <- calc_pp(auroc_dem, "education")
```

Bind all pp/contrast tibbles and save
```{r}
list(sex[[1]], race[[1]], income[[1]], age[[1]], education[[1]]) |> 
  bind_rows() |> 
  write_csv(here::here(path_models, "posteriors_fairness_v10_full.csv"))

list(sex[[2]], race[[2]], income[[2]], age[[2]], education[[2]]) |> 
  bind_rows() |> 
  write_csv(here::here(path_models, "pp_fairness_full.csv"))

list(sex[[3]], race[[3]], income[[3]], age[[3]], education[[3]]) |> 
  bind_rows() |>
  write_csv(here::here(path_models, "pp_fairness_contrast_full.csv"))
```

## LLs

> CP note: is it helpful or useful to get LL by demographic group? my understanding is that LL is calculated for each observation, so I'm not sure if this is 

```{r}
get_ll_by_dem <- function(df, var) {
  # df should be dataframe with demographic features and predictions
  # var should be character or factor variable in df
  
  var_tmp <- df |> 
      pull(get(var))
   
  var_levels <- 
     if (class(var_tmp) != "factor") {
       levels(as.factor(var_tmp))
     } else levels(var_tmp)
  
  # create empty tibble with split nums
  ll <- tibble(outer_split_num = unique(df$outer_split_num)) |> 
    arrange(outer_split_num) 
    
  for (level in var_levels) {
     ll_level <- df |>
       filter(get(var) == level) |> 
       nest(.by = outer_split_num, .key = "preds") |> 
       mutate(ll = map(preds, \(preds) mn_log_loss(preds, prob_raw, 
                                             truth = label))) |> 
       select(-preds) |> 
       unnest(ll) |> 
       select(-c(.estimator, .metric)) |> 
       rename(!!level:=.estimate)
     
     ll <- ll |> 
       left_join(ll_level, by = "outer_split_num")
  }
  
  return(ll)
}
```

get LLs
```{r}
ll_dem <- get_ll_by_dem(df = feat_preds, var = "demo_sex") |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "demo_race"), by = "outer_split_num") |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "income_split"), by = "outer_split_num") |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "age_split"), by = "outer_split_num")  |> 
  left_join(get_ll_by_dem(df = feat_preds, var = "demo_educ_comb"), by = "outer_split_num")  
```

Save auroc_dem
```{r}
#auroc_dem |> 
#  write_csv(here::here(path_models, "fairness_auroc_6_x_5_1_x_5_nested_v10_full.csv"))  
```
