---
title: "Model probabilities"
author: "John Curtin, edited by Claire for study_gps"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  version: "v25"
  cv: "nested_2_x_5_6_x_5"
  model: "full_v_baseline"
  batch: "full"
editor_options: 
  chunk_output_type: console
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

options(conflicts.policy = "depends.ok")
library(probably)

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_input <- format_path("risk/models/gps")

version <- params$version
cv <- params$cv
model <- params$model
batch <- params$batch
```

```{r}
preds <- read_csv(here::here(path_input, str_c("outer_preds_", 
                                            version, "_", 
                                            cv, "_", batch, ".csv")))

metrics <- read_csv(here::here(path_input, str_c("outer_metrics_", 
                                           version, "_", 
                                           cv, "_", batch, ".csv")))

```

## Basic histogram and calibration plot

Get lapse prevalence and log loss in each fold.
```{r}
text_labels <- preds |> 
  group_by(outer_split_num) |> 
  mutate(prev = mean(label == "lapse"),
         log_loss_raw = mn_log_loss_vec(truth = factor(label,
                                                   levels = c("lapse",
                                                              "no lapse")),
                                    estimate = prob_raw)) |> 
  ungroup() |> 
  merge(metrics, by = "outer_split_num") |> 
  select(outer_split_num, prev, roc_auc, log_loss_raw, hp1, hp2, hp3) |>
  distinct(outer_split_num, .keep_all = TRUE)

text_labels <- text_labels |>
  mutate(
    label = paste0(
      "prevalence: ", round(prev, 3),
      "\nlog loss: ", round(log_loss_raw, 3),
      "\nhp1: ", hp1,
      "\nhp2: ", hp2,
      "\nhp3: ", hp3,
      "\nauroc: ", round(roc_auc, 3)
    ))
```
 
Create faceted histograms.
```{r}
preds |> 
  ggplot(aes(x = prob_raw)) +
    geom_histogram(binwidth = .01) +
    labs(title = "Histogram of Raw Probabilities") + 
    scale_x_continuous(breaks = seq(0, 1, .25)) +
    coord_cartesian(xlim = c(0, 1)) +
    facet_wrap(~ outer_split_num) +
  geom_text(
  data = text_labels,
  aes(x = Inf, y = Inf, label = label),
  hjust = 1.05, vjust = 1.1, size = 2.5) +
  theme_light()
```

Calibration plot
```{r}
bin_width = 0.10
cal_raw <- preds |>
  mutate(
    prob = prob_raw,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "raw"
  )


cal_raw |> 
  group_by(bins, probs)  |>
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse, color = probs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_line() +
    geom_point() +
    labs(x = "Predicted Lapse Probability (bin mid-point)",
         y = "Observed Lapse Probability",
         title = "Calibration (Raw Probabilities)") +
    scale_x_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
  theme_minimal()

```

Create a plot of probabilities per fold, so we can see the variability within folds.
```{r}
bin_width = 0.10
cal_raw <- preds |>
  mutate(
    prob = prob_raw,
    bins = cut(prob, breaks = seq(0, 1, bin_width)),
    lapse = if_else(label == "lapse", 1, 0),
    probs = "raw"
  )


cal_raw |> 
  group_by(outer_split_num, bins, probs)  |>
  summarize(mean_lapse = mean(lapse),
            n = n(),
            .groups = "drop") |>
  filter(n >= 50) |> 
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_line(color = "darkgreen") +
    geom_point(color = "darkgreen") +
    labs(x = "Predicted Lapse Probability (bin mid-point)",
         y = "Observed Lapse Probability",
         title = "Raw Probabilities by Fold") +
    scale_x_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
  facet_wrap(~ outer_split_num) +
  theme_classic()

```

Check how many observations are in a given bin -- this might help account for the instability we see across folds. This is collapsing across folds...
```{r}
cal_raw |>
  group_by(bins, probs) |>
  summarise(
    mean_pred = mean(prob),
    mean_lapse = mean(lapse),
    n = n(),
    .groups = "drop"
  ) |> 
  view()
```

...but we can also look at this grouped by folds to get better insight into our plots. It looks like part of the issue with looking at our calibration plots is that some of our bins have very few observations.
```{r}
cal_raw |>
  group_by(bins, probs, outer_split_num) |>
  summarise(
    mean_pred = mean(prob),
    mean_lapse = mean(lapse),
    n = n(),
    .groups = "drop"
  ) |>
  view()
```

We can also print a table of the absolute difference between predicted and true values of lapse for each probability type.
```{r}
cal_raw |>
  group_by(outer_split_num, probs, bins) |>
  summarize(
    mean_lapse = mean(lapse),
    mean_pred = mean(prob),
    .groups = "drop"
  ) |>
  mutate(
    calibration_error = abs(mean_pred - mean_lapse)
  ) |>
  group_by(probs) |>
  summarize(
    mean_cal_error = mean(calibration_error),
    median_cal_error = median(calibration_error),
    max_cal_error = max(calibration_error)
  )
```

## Brier scores

```{r}
brier_scores <- preds |>
  pivot_longer(
    cols = prob_raw,
    names_to = "method",
    values_to = ".pred_lapse") |>
  mutate(label = factor(label,
                        levels = c("lapse", "no lapse"))) |> 
  group_by(outer_split_num,
           method) |>
  brier_class(label, .pred_lapse)
```

```{r}
brier_scores |>
  group_by(method) |>
  summarise(
    mean_brier = mean(.estimate),
    sd_brier = sd(.estimate),
    median_brier = median(.estimate),
    min_brier = min(.estimate),
    max_brier = max(.estimate)
  )
```