---
title: "Reprex: Investigating `uniroot()` failure in `betacal:beta_calibration()`"
author: "Claire Punturieri"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  study: "gps"
  version: "v9"
  cv: "kfold_5_x_5"
  algorithms: "xgboost"   # "all" or name of specific algorithm
  model: "context_movement_weather"
editor_options: 
  chunk_output_type: console
---

### Note 

This shows how uniroot() can fail during beta calibration when the calibration function doesn't change sign over the interval (requirement of uniroot()).

This was created to explore why cal_estimate_beta() is failing in mak_local_preds.qmd.

Error running into:
> Error in uniroot(function(mh) b * log(1 - mh) - a * log(mh) - inter, c(1e-16,  : 
  f() values at end points not of opposite sign
  
Error occurs in cal_estimate_beta() which uses betacal::beta_calibration().

### Set Up Environment

Function conflicts
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

options(conflicts.policy = "depends.ok")
library(probably)

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
source("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
source("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("https://github.com/jjcurtin/lab_support/blob/main/chtc/static_files/fun_chtc.R?raw=true")

path_input <- format_path(str_c("risk/chtc/gps"))
path_models <- format_path(str_c("risk/models/gps"))
```

### Reprex set-up

Load in dataset.
```{r}
config_best <- 
  read_csv(here::here(path_models, 
                      str_c("best_config_",
                            params$version, "_",
                            params$cv, "_",
                            params$model, ".csv"))) |>
  mutate(n_folds = n(),
         algorithm = "xgboost") |> 
  slice(1) |> 
  select(feature_set, hp1, hp2, hp3, resample, n_folds, split_num, algorithm) |>
  glimpse()
```

Set up splits.
```{r} 
batch_names <- list.dirs(path_input, full.names = FALSE, recursive = FALSE)
batch_name <- batch_names[str_detect(batch_names, params$version)]
path_batch <- here::here(path_input, batch_name)
source(here::here(path_batch, "input", "training_controls.R"))
          
chunks <- str_split_fixed(data_trn, "\\.", n = Inf) # parse name from extensions
if (length(chunks) == 2) {
  fn <- str_c("data_trn.", chunks[[2]])
} else {
  fn <- str_c("data_trn.", chunks[[2]], ".", chunks[[3]])
}
  
# open based on file type
if (str_detect(fn, "csv")) {
  d <- read_csv(here::here(path_batch, "input", fn), show_col_types = FALSE) 
} else {
  d <- read_rds(here::here(path_batch, "input", fn))
}

d <- d |> 
  arrange(subid, dttm_label) |> 
  format_data()

splits <- d |> 
  make_splits(cv_resample_type, cv_resample, cv_outer_resample, 
              cv_inner_resample, cv_group, cv_strat = cv_strat,
              the_seed = seed_splits)
```

Define split_num for testing. Known failure at split_num = 3.
```{r}
split_num <- 3
```

Simplified recipe set-up.
```{r}
d_in <- splits$splits[[split_num]] |> training()
d_out <- splits$splits[[split_num]] |> testing()

rec <- build_recipe(d = d_in, config = config_best)

set.seed(2468)

cal_split <- d_in |> 
  group_initial_split(group = all_of(cv_group), strata = strat, prop = 3/4)

d_cal_in <- training(cal_split)
d_cal_out <- testing(cal_split)

rec_cal_prep <- rec |> prep(training = d_cal_in)

feat_cal_in <- rec_cal_prep |> bake(new_data = NULL)
feat_cal_out <- rec_cal_prep |> bake(new_data = d_cal_out)

# Train model on calibration training fold
model_cal <- fit_best_model(config_best, feat = feat_cal_in, "classification")
```

### Run beta_calibration()

First run predict().
```{r}
beta <- predict(model_cal, feat_cal_out,
                type = "prob")
```

Manually set p and y (inputs to betacal::beta_calibration()).
```{r}
p <- beta$.pred_lapse
y <- feat_cal_out$y
```

Change y from factor to numeric (from cal_estimate_beta(), line 188).
```{r}
lvls <- levels(y)
y <- ifelse(y == lvls[1], 1, 0)
```

Run betacal::beta_calibration() with "abm" parameter (default parameter and what we are currently using).
```{r}
p <- pmax(1e-16, pmin(p, 1-1e-16))

d <- data.frame(y)
d$lp <- log(p)
d$l1p <- -log(1-p)

fit <- glm(y~lp+l1p,family=binomial(link='logit'),data=d)

a <- as.numeric(fit$coefficients['lp'])
b <- as.numeric(fit$coefficients['l1p'])
if (a < 0){
  fit <- glm(y~l1p,family=binomial(link='logit'),data=d)
  a <- 0
  b <- as.numeric(fit$coefficients['l1p'])
  } else if (b < 0){
    fit <- glm(y~lp,family=binomial(link='logit'),data=d)
    a <- as.numeric(fit$coefficients['lp'])
    b <- 0
  }

inter <- as.numeric(fit$coefficients['(Intercept)'])
    
#m <- uniroot(function(mh) b*log(1-mh)-a*log(mh)-inter,c(1e-16,1-1e-16))$root

#calibration <- list("map" = c(a,b,m), "model" = fit, "parameters" = parameters)

## format as trycatch for rendering
tryCatch({
  m <- uniroot(function(mh) b*log(1 - mh) - a*log(mh) - inter,
               c(1e-16, 1 - 1e-16))$root
  
  calibration <- list("map" = c(a, b, m), "model" = fit, "parameters" = parameters)
  print("uniroot() succeeded. Calibration map calculated.")
}, error = function(e) {
  message("uniroot() failed: ", e$message)
  calibration <- NULL
})
```

Here is a demonstration of why uniroot() fails in some cases. uniroot() will fail when the function does not change sign over the interval, which it expects. This is because it wants to find the root of a continuous function over an interval (AKA where f(x) = 0). If it doesn't pass 0, it dies on us!
```{r}
# write out the function separately
f <- function(mh) b*log(1-mh)-a*log(mh)-inter

# pull out the endpoints resulting from the above abm values
left <- f(1e-16)
right <- f(1-1e-16)

cat("f(1e-16):", left, "\n")
cat("f(1-1e-16):", right, "\n")

# check if signs are opposite
if (sign(left) == sign(right)) {
  message("Function does not change sign over interval. uniroot() would fail.")
} else {
  message("Function does change sign over interval. uniroot() would work -- yay!")
}
```

## Potential solutions

1) Try specifying a different parameter set instead of abm. uniroot() is only used in abm (which is "full"/"true" beta calibration and is the most flexible), so we could consider using an alternative like am or ab. However, these options might not do as good of a job as abm at correcting miscalibration.

2) Modify the beta_calibration function to skip using uniroot() in cases where the function does not change sign over interval. **Probably not a good idea.**